# Missing values

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r, echo=FALSE} 
library(tidyverse)
library(patchwork)
library(magrittr)
``` 

## Load Dataset
```{r}
load("~/Documents/MSDS/EDAV/Final Project Data/DS0001/33661-0001-Data.rda")
df <- da33661.0001
```

## Evaluating Missing Values

### Defining a Function to Plot Missing Value Patterns: miss_pat()

miss_pat takes in two values, (1) data: the dataframe of interest and (2) style: specifying 'count' or 'percent' as the style of graph you are interested in displaying. 

```{r}
miss_pat <- function(data, style) {
  missing_patterns_tidy <- data.frame(is.na(data)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup() %>%
    rownames_to_column("pattern id") %>%
    pivot_longer(cols = !c(`pattern id`, count), names_to = "feature", values_to = "value") %>%
    group_by(feature) %>%
    mutate(freq = sum(count * ifelse(value == TRUE, 1, 0))) %>%
    ungroup()
  
  missing_patterns_tidy_complete <- missing_patterns_tidy %>%
    group_by(`pattern id`) %>%
    mutate(num = sum(ifelse(value == TRUE, 1, 0))) %>%
    group_by(`pattern id`) %>%
    mutate(missing = ifelse(sum(num) == 0, "complete", "missing")) %>%
    mutate(value = ifelse(missing == "complete", "complete", as.character(value))) %>%
    mutate(complete = ifelse(missing == "complete" & feature == feature[1], "complete cases", NA))
  
  # setup data frame for top bar chart, missing per variable/feature
  missing_patterns_tidy_1 <- missing_patterns_tidy %>%
    select(feature, freq) %>%
    unique()
  
  missing_patterns_tidy_1 <- missing_patterns_tidy_1 %>%
    mutate(percent = sum(missing_patterns_tidy_1$freq)) %>%
    mutate(percent = 100 * (freq/percent))

    
  # create data frame for side bar chart with only pattern id and count information, excluding repeat rows; compute count percentages
  missing_patterns_tidy_2 <- missing_patterns_tidy_complete %>%
    select(`pattern id`, count, missing) %>%
    unique()
    
  missing_patterns_tidy_2 <- missing_patterns_tidy_2 %>%
    mutate(percent = sum(missing_patterns_tidy_2$count)) %>%
    mutate(percent = 100 * (count/percent))
  
  
  # conditional next steps based on count or percent
  # if user wants counts to be plotted
  if (style == 'count') {
    # set y variable and and scale according to frequency choice
    option_1 = missing_patterns_tidy_1$freq
    upper_y_1 = max(option_1)
    label_y_1 = "Number of\nMissing Rows"
    
    option_2 = missing_patterns_tidy_2$count
    upper_y_2 = max(option_2)
    label_y_2 = "Row Count"
  }
  
  # if user wants percents to be plotted
  if (style == 'percent') {
    # set y variable and and scale according to percent choice
    option_1 = missing_patterns_tidy_1$percent
    upper_y_1 = sum(option_1)
    label_y_1 = "Percentage of\nMissing Rows"
    
    option_2 = missing_patterns_tidy_2$percent
    upper_y_2 = 100
    label_y_2 = "Percentage\nof Rows"
  }
  
  
  # create center pattern plot
  plot_center <- ggplot(missing_patterns_tidy_complete, aes(x = fct_reorder(as.factor(feature), -freq), y = fct_rev(as.factor(as.numeric(`pattern id`))),
                                                      fill = value, alpha = 0.7)) +
  geom_tile(color = "white", alpha = .7, show.legend = FALSE) +
  geom_text(aes(label = complete, x = 4), show.legend = FALSE) +
  scale_fill_manual(values = c("darkgray", "lightgray", "mediumpurple1")) +
  xlab("Variable") + 
  ylab("Missing Pattern") +
  theme_classic(8)

  
  # top bar chart: frequency or percentage of missing row by car feature
  plot_top <- ggplot(missing_patterns_tidy_1, aes(x = fct_reorder(as.factor(feature), -option_1), y = option_1)) +
    geom_col(fill = "royalblue1", alpha = 0.6) +
    ylim(0,upper_y_1) +
    ggtitle("Missing Value Patterns") +
    xlab("") +
    ylab(label_y_1) +
    theme_classic(8)
  
  # side bar chart: frequency or percentage of missing row by pattern id
  plot_side <- ggplot(missing_patterns_tidy_2, aes(x = fct_rev(as.factor(as.numeric(`pattern id`))), 
                                                   y = option_2, fill = missing, alpha = .9)) +
    geom_col(show.legend = FALSE) +
    scale_fill_manual(values = c("blue", "royalblue1")) +
    coord_flip() +
    ylim(0,upper_y_2) +
    xlab("") +
    ylab(label_y_2) +
    theme_classic(8)
  
  #combine using patchwork
  (plot_top + plot_spacer() + plot_layout(widths = c(1.6,.75))) /
  (plot_center + plot_side + plot_layout(widths = c(2,.75))) +
  plot_layout(heights = c(.75, 2))
    
}

```

### Visualizing Missing Patterns

Since our dataset it quite large (795, 590), we split up our visualizations per 100 columns, so patterns are still interpretable. However, the downside to this approach is that we lose sight of overarching patterns of missing data across all cases, what it does show is amongst a 100 column sample, which columns or attributes are most commonly missing. 

```{r}
#abbreviate column names for graphing purposes
names(df) <- abbreviate(names(df))

miss_pat((df)[500:580], 'percent')
miss_pat((df)[400:500], 'percent')
miss_pat((df)[300:400], 'percent')
miss_pat((df)[200:300], 'percent')
miss_pat((df)[100:200], 'percent')
miss_pat((df)[0:100], 'percent')

```

### Observations 

* 








