# Missing values

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r, echo=FALSE} 
library(tidyverse)
library(patchwork)
library(magrittr)
``` 

```{r}
#load dataset
load("~/Documents/MSDS/EDAV/Final Project Data/DS0001/33661-0001-Data.rda")
df <- da33661.0001

#abbreviate column names for graphing purposes
names(df) <- abbreviate(names(df))
```

## Defining a Function to Plot Missing Value Patterns: miss_pat()

`miss_pat` takes in two values, 

(1) `data`: the dataframe of interest and 

(2) `style`: specifying 'count' or 'percent' as the style of graph you are interested in displaying. 

```{r, function}
miss_pat <- function(data, style) {
  missing_patterns_tidy <- data.frame(is.na(data)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup() %>%
    rownames_to_column("pattern id") %>%
    pivot_longer(cols = !c(`pattern id`, count), names_to = "feature", values_to = "value") %>%
    group_by(feature) %>%
    mutate(freq = sum(count * ifelse(value == TRUE, 1, 0))) %>%
    ungroup() 

  # only keep values where count of pattern is at least 5 for visualization purposes
  missing_patterns_tidy <- missing_patterns_tidy[missing_patterns_tidy$count > 5, ]
  
  missing_patterns_tidy_complete <- missing_patterns_tidy %>%
    group_by(`pattern id`) %>%
    mutate(num = sum(ifelse(value == TRUE, 1, 0))) %>%
    group_by(`pattern id`) %>%
    mutate(missing = ifelse(sum(num) == 0, "complete", "missing")) %>%
    mutate(value = ifelse(missing == "complete", "complete", as.character(value))) %>%
    mutate(complete = ifelse(missing == "complete" & feature == feature[1], "complete cases", NA))
  
  # setup data frame for top bar chart, missing per variable/feature
  missing_patterns_tidy_1 <- missing_patterns_tidy %>%
    select(feature, freq) %>%
    unique()
  
  missing_patterns_tidy_1 <- missing_patterns_tidy_1 %>%
    mutate(percent = sum(missing_patterns_tidy_1$freq)) %>%
    mutate(percent = 100 * (freq/percent))

    
  # create data frame for side bar chart with only pattern id and count information, excluding repeat rows; compute count percentages
  missing_patterns_tidy_2 <- missing_patterns_tidy_complete %>%
    select(`pattern id`, count, missing) %>%
    unique()
    
  missing_patterns_tidy_2 <- missing_patterns_tidy_2 %>%
    mutate(percent = sum(missing_patterns_tidy_2$count)) %>%
    mutate(percent = 100 * (count/percent))
  
  
  # conditional next steps based on count or percent
  # if user wants counts to be plotted
  if (style == 'count') {
    # set y variable and and scale according to frequency choice
    option_1 = missing_patterns_tidy_1$freq
    upper_y_1 = max(option_1)
    label_y_1 = "Number of\nMissing Rows"
    
    option_2 = missing_patterns_tidy_2$count
    upper_y_2 = max(option_2)
    label_y_2 = "Row Count"
  }
  
  # if user wants percents to be plotted
  if (style == 'percent') {
    # set y variable and and scale according to percent choice
    option_1 = missing_patterns_tidy_1$percent
    upper_y_1 = 100
    label_y_1 = "Percentage of\nMissing Rows"
    
    option_2 = missing_patterns_tidy_2$percent
    upper_y_2 = 100
    label_y_2 = "Percentage\nof Rows"
  }
  
  
  # create center pattern plot
  plot_center <- ggplot(missing_patterns_tidy_complete, aes(x = fct_reorder(as.factor(feature), -freq), 
                                                            y = fct_rev(as.factor(as.numeric(`pattern id`))),
                                                      fill = value, alpha = 0.7)) +
  geom_tile(color = "white", alpha = .7, show.legend = FALSE) +
  geom_text(aes(label = complete, x = 4), show.legend = FALSE) +
  scale_fill_manual(values = c("darkgray", "lightgray", "mediumpurple1")) +
  xlab("Variable") + 
  ylab("Missing Pattern") +
  theme_classic(8)

  
  # top bar chart: frequency or percentage of missing row by car feature
  plot_top <- ggplot(missing_patterns_tidy_1, aes(x = fct_reorder(as.factor(feature), -option_1), y = option_1)) +
    geom_col(fill = "royalblue1", alpha = 0.6) +
    ylim(0,upper_y_1) +
    ggtitle("Missing Value Patterns") +
    xlab("") +
    ylab(label_y_1) +
    theme_classic(8)
  
  # side bar chart: frequency or percentage of missing row by pattern id
  plot_side <- ggplot(missing_patterns_tidy_2, aes(x = fct_rev(as.factor(as.numeric(`pattern id`))), 
                                                   y = option_2, fill = missing, alpha = .9)) +
    geom_col(show.legend = FALSE) +
    scale_fill_manual(values = c("blue", "royalblue1")) +
    coord_flip() +
    ylim(0,upper_y_2) +
    xlab("") +
    ylab(label_y_2) +
    theme_classic(8)
  
  #combine using patchwork
  (plot_top + plot_spacer() + plot_layout(widths = c(1.6,.75))) /
  (plot_center + plot_side + plot_layout(widths = c(2,.75))) +
  plot_layout(heights = c(.75, 2))
    
}

```

## Visualizing Missing Patterns

Since our dataset is quite large we only included missing patterns that had a count of at least 5. 

```{r fig12}
miss_pat(df, 'count')
miss_pat(df, 'percent')
```

### Observations 

* The most common pattern is of complete cases, which is a great sign for our analysis! About 65% of rows contain no missing variables. 

* For patterns 2, 4, 5 & 7 - the pattern is fairly similar across and spans seemingly similar sets of variables. 

* Patterns 3 & 6 are also similar in the types of missing data. 

## A Deeper Look into the Missing Data 

```{r fig3, fig.height = 8, fig.width = 6, fig.align = "center"}

missing_patterns_tidy <- data.frame(is.na(df)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup() %>%
  rownames_to_column("pattern id") %>%
  pivot_longer(cols = !c(`pattern id`, count), names_to = "feature", values_to = "value") %>%
  group_by(feature) %>%
  mutate(freq = sum(count * ifelse(value == TRUE, 1, 0))) %>%
  ungroup() 

# only keep values where count of pattern is at least 5 for visualization purposes
missing_patterns_tidy <- missing_patterns_tidy[missing_patterns_tidy$count > 5, ]
  
missing_patterns_tidy_complete <- missing_patterns_tidy %>%
  group_by(`pattern id`) %>%
  mutate(num = sum(ifelse(value == TRUE, 1, 0))) %>%
  group_by(`pattern id`) %>%
  mutate(missing = ifelse(sum(num) == 0, "complete", "missing")) %>%
  mutate(value = ifelse(missing == "complete", "complete", as.character(value))) %>%
  mutate(complete = ifelse(missing == "complete" & feature == feature[1], "complete cases", NA))

ggplot((missing_patterns_tidy_complete[missing_patterns_tidy_complete$freq > 25,] ), aes(x = fct_reorder(as.factor(feature), -freq), 
                                                            y = (as.factor(as.numeric(`pattern id`))),
                                                      fill = value, alpha = 0.7)) +
  geom_tile(color = "white", alpha = .7, show.legend = FALSE) +
  geom_text(aes(label = complete, x = 4), show.legend = FALSE) +
  scale_fill_manual(values = c("darkgray", "lightgray", "mediumpurple1")) +
  coord_flip() +
  xlab("Variable") + 
  ylab("Missing Pattern") +
  ggtitle("Zoom in to Missing Patterns (Flipped)") +
  theme_classic(10)

```

To take a deeper look into the data that is missing, we took a look directly at the 'missing' areas that were shown above. For the purposes of analysis we also flipped the chart in order to better read the column headers that constitute these missing data patterns. Note that the chart now increases in frequency of 'missing' from left to right on the x-axis per pattern. 

### Observations: 

* Based on inspection, we can see that the common questions of the questionnaire that have missing values are those in the following categories: 
  * MPSQ: Reasons for Participation in Sport
  
  * CESD: Mental Health Assessment Section
  
  * SRI: Sexual Activity Questions

* The two variables that are missing across 4 patterns each are: 

  (1) SRI02 - LAST TIME HAD SEX: USED OTHER BIRTH CONTROL

  (2) PSME - RECEIVED MEDIA ATTENTION 

* For patterns 3 and 6 the common areas of missing variables are from:

  * CESD: the Mental Health Assessment section 
  
  * SRI: the Sexual Activity Section 

* The variable missing most often is TEOSQ13. 

#### Discussion 

* It is understandable that some college students would prefer not to disclose their sexual behavior, so this pattern of missing values seems non-random. 

* Mental health likewise is topic that is often glossed over, and given the survey was conducted in 2006 mental health was an even less discussed topic than it is today, it is not surprising that some of these questions were left unanswered by students. 

* The MSPQ variables are part of the following question: "People have many different reasons for participating in sports. Below are listed some of those possible reasons. How important is each of these reasons for you?" and each variable accounts for a different answer to the question which the student could answer on a score from 1 to 5. A missing variable indicates that they did not answer the question at all. This could be because they do not participate in sports. 

* Variable TEOSQ13 is a question within the "Feelings about Athletic Success" section where the question was: "People have different ideas about what athletic "success" means, other than simply winning or losing a competition. Think about the times when you have felt most successful in sport. How well does each of the following statements describe your feelings? I feel most successful in sport when . . . I do my very best." There were 13 variations on this question that provided different answers to select from in the questionnaire and all 13 of the TEOSQ variables show up in the missing values top list. Similar to the MSPQ variables, the responses to the question was a score between 1 to 5. We believe that missing values to this question are also due to students who did not participate in sports. 



